---
title: "Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method"
editor: visual
---

# Overview

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable).

In this hands on exercise, we will learn how to build ***hedonic pricing*** model by using the GWR method. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

# The Data

In this exercide we will use two data sets. They are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

# Getting Started 

The following R packages will be installed into R environment.

-   **olsrr:** building OLS and performing diagnostics tests

-   **GWmodel**: calibrating geographical weighted family of models

-   **coorplot:** multivariate data visualisation and analysis

-   **sf:** spatial data handling

-   **tidyverse**, especially **readr**, **ggplot2** and **dplyr: Attribute data handing**

-   **tmap:** choropleth mapping

The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

# A note about GWmodel 

This package provides a collection of localised spatial statistical methods, namely:

1.   GW summary statistics,
2.  GW principal components analysis,
3.  GW discriminant analysis and
4.  various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms.

Commonly, outputs or parameters of the GWmodel are mapped to provide a useful exploratory tool, which can often precede (and direct) a more traditional or sophisticated statistical analysis.

# Geospatial Data Wrangling 

## Importing geospatial data 

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

Learning from above code chunk:

**mpsz** is the imported shapefile and it is simple feature object.

## Updating CRS Information

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

Using `st_crs()` of **sf** package, we can verify the projection of the newly transformed *mpsz_svy21*

The code chunk below will be used to varify the newly transformed *mpsz_svy21*.

```{r}
st_crs(mpsz_svy21)
```

Next, we will reveal the extent of *mpsz_svy21* by using `st_bbox()` of sf package.

```{r}
st_bbox(mpsz_svy21) #view extent
```

# Importing Aspatial Data Wrangling 

The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

The codes chunks below uses `glimpse()` to display the data structure of will do the job.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

Next, `summary()` of base R is used to display the summary statistics of *cond_resale* tibble data frame.

```{r}
summary(condo_resale)
```

## Converting aspatial data frame into a sf object

The code chunk below converts condo_resale data frame into a simple feature data frame by using `st_as_sf()` of **sf** packages.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

```{r}
head(condo_resale.sf)  #list contents 
```

**Note: The ouput is in point feature data frame.**

# Exploratory Data Analysis (EDA)

In the section, we will learn how to use statistical graphics functions of **ggplot2** package to perform EDA.

## Stattistical Graphs 

Using the code chunk below, we will plot the distribution of Selling_Price

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

**The figure above reveals a right skewed distribution. This means that more condominium units were transacted at relative lower prices.**

The code chunk below is used to derive a new variable called *LOG_SELLING_PRICE* by using a log transformation on the variable *SELLING_PRICE*. It is performed using `mutate()` of **dplyr** package.

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

Using the code chunk below, Log_selling_Price is plotted

```{r}
ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

**Notice that the distribution is relatively less skewed after the transformation.**

## Multiple Histogram Plots distribution of variables

Using `ggarrange()` of [**ggpubr**](https://cran.r-project.org/web/packages/ggpubr/) package, multiple histograms will be plotted.
The code chunk below is used to create 12 histograms. Then, `ggarrange()` is used to organised these histogram into a 3 columns by 4 rows small multiple plot.

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, 
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

## Drawing Statistical Point Map 

In this section, using tmap, we will reveal the geospatial distribution condominium resale prices in Singapore.

```{r}
tmap_mode("view")  #interactive mode
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

`set.zoom.limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

Before moving on to the next section, the code below will be used to turn R display into `plot` mode.

```{r}
tmap_mode("plot")  #static mode
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

# Hedonic Pricing Modelling in R









 









# 




