---
title: "Geospatial Analytics for Social Good"
editor: visual
---

# Background

Water is an essential resource, providing healthy environment, a sustainable economy, reducing poverty and also ensuring peace and security. Importance of accessible and clear water is very critical, already 40% of the population doesn't have access to safe drinkable water. The scarcity of water leads to imbalance in the ecosystem. There will be food scarcity as agricultural department is dependent on world's accessible freshwater.

Scarcity of fresh water is so adverse that 80% of illness in developing countries are because of lack of sanitation and clean water. Providing clean water in rural communities of the developing countries is still a major issue in many countries globally, especially countries in Africa continent.

To address this issue a projected has been initiated with aim to collect water point related data from rural areas at the water point or small water scheme level and share the data via WPdx Data Repository, a cloud-based data library.

# Objective

To address complex problems faced by society using geospatial analysis. Using appropriate global and local measures of spatial Association techniques, the spatial patterns of Not Functional water points are revealed. In this exercise, Nigeria will be used as a study point country.

# The Data

## Aspatial Data

The data from [WPdx Global Data Repositories](https://www.waterpointdata.org/access-data/) will be used, WPdx+ data set.

## Geospatial Data

Nigeria Level-2 Administrative Boundary (also known as Local Government Area) polygon features GIS data will be used downloaded from [geoBoundaries](https://www.geoboundaries.org/).

# The Task

## Data Wrangling

Following tasks will be performed in geospatial analysis:

-   Using appropriate sf method, importing the shapefile into R and save it in a simple feature data frame format.

-   Using appropriate tidyr and dplyr methods, proportion of functional and non-functional water point at LGA level is derived.

-   Combining the geospatial and aspatial data frame into simple feature data frame.

-   Performing outliers/clusters analysis by using appropriate local measures of spatial association methods.

-   Performing hotspot areas analysis by using appropriate local measures of spatial association methods.

## Thematic Mapping

-   Maps are plotted to show the spatial distribution of functional and non-functional water point rate at LGA level by using appropriate thematic mapping technique provided by tmap package.

## Analytical Mapping

-   Hotspot areas and outliers/clusters maps of functional and non0functional water point rate at LGA level by using appropriate thematic mapping technique provided by tmap package.

# Getting Started

## Setting the Analytical Tools

Installing and launching R packages into R environment

```{R}
pacman::p_load(sf, spdep, tmap, tidyverse,funModeling)
```

## Import shapefile into R environment

### Importing water point geospatial

Using, st_read() of **sf** package, Nigeria Level-2 Administrative Boundary shapefile is imported into R environment.

```{r}
#| eval: false

wp <- st_read(dsn = "Geodata",
              layer = "geo_export",
              crs = 4326) %>%
  filter(clean_coun == "Nigeria")
```

Things to learn from the code chunk above:

\*\*st_read()\*\* of sf package is used to import geo_export shapefile into R environment and save the imported geospatial data into simple feature data table.

\*\*filter()\*\* of dplyr package is used to extract water point records of Nigeria. In the code chunk below, "write_rds()" of reader package is used to save the extracted sf table .....

Next, \`write_rds()\` of readr package is used to save the extracted sf data table (i.e. wp) into an output file in rds data format. The output file is called \*wp_nga.rds\* and it is saved in \*geodata\* sub-folder.

```{R}
#| eval: false
write_rds(wp, "Geodata/wp_nga.rds")
```

### Importing Nigeria LGA boundary data

Now, we are going to import the LGA boundary data into R environment by using the code chunk below.

```{R}
#| eval: false
nga <- st_read(dsn = "Geodata",
               layer= "geoBoundaries-NGA-ADM2",
               crs = 4326)
```

Thing to learn from the code chunk above.

\- \`st_read()\` of \*\*sf\*\* package is used to import \*geoBoundaries-NGA-ADM2\* shapefile into R environment and save the imported geospatial data into simple feature data table.

# Data Wrangling

## Recoding the NA values into string

In the code chunk below, \`replace_na()\` is used to recode all the \*NA\* values in \*status_cle\* field into \*Unknown\*.

```{R}
#| eval: false
wp_nga <- read_rds("geodata/wp_nga.rds") %>%
  mutate(status_cle = replace_na(status_cle, "Unknown"))
  
```

## Exploratory Data Analysis

In the code chunk below, `freq()` of **funModeling** package is used to display the distribution of *status_cle* field in *wp_nga*.

```{r}
#| eval: false
freq(data=wp_nga, 
     input = 'status_cle')
```

# Extracting Water Points

By using classes in *status_cle* field, extracting water points records.

## Extracting functional water point

Using, `filter()` of dplyr is used to select functional water points.

```{r}
#| eval: false
wpt_functional <- wp_nga %>%
  filter(status_cle %in%
           c("Functional", 
             "Functional but not in use",
             "Functional but needs repair"))
```

```{r}
#| eval: false
freq(data=wpt_functional, 
     input = 'status_cle')
```

## Extracting non-functional water points

Using, `filter()` of dplyr is used to select non-functional water points.

```{r}
#| eval: false
wpt_nonfunctional <- wp_nga %>%
  filter(status_cle %in%
           c("Abandoned/Decommissioned", 
             "Abandoned",
             "Non-Functional",
             "Non functional due to dry season",
             "Non-Functional due to dry season"))
```

```{r}
#| eval: false
freq(data=wpt_nonfunctional, 
     input = 'status_cle')
```

## Extracting water point with unknown class

Using, `filter()` of dplyr is used to select water points with unknown status.

```{r}
#| eval: false
wpt_unknown <- wp_nga %>%
  filter(status_cle == "Unknown")
```

# Performing Point-in Polygon Count

```{r}
#| eval: false
nga_wp <- nga %>% 
  mutate(`total wpt` = lengths(
    st_intersects(nga, wp_nga))) %>%
  mutate(`wpt functional` = lengths(
    st_intersects(nga, wpt_functional))) %>%
  mutate(`wpt non-functional` = lengths(
    st_intersects(nga, wpt_nonfunctional))) %>%
  mutate(`wpt unknown` = lengths(
    st_intersects(nga, wpt_unknown)))
```

# Saving the Analytical Data Table

Learnings from code chunks:

-   `mutate()` of **dplyr** package is used to derive two fields namely *pct_functional* and *pct_non-functional*.

-   to keep the file size small, `select()` of **dplyr** is used to retain only field 3,4,9,10, 18,19,20,21,22,and 23.

```{r}
#| eval: false
nga_wp <- nga_wp %>%
  mutate(pct_functional = `wpt functional`/`total wpt`) %>%
  mutate(`pct_non-functional` = `wpt non-functional`/`total wpt`)
```

```{r}
#| eval: false
write_rds(nga_wp, "Geodata/nga_wp.rds")
```

# Visualising the spatial distribution of water points

```{r}
nga_wp <- read_rds("Geodata/nga_wp.rds")
#| fig-width: 14
#| fig:height: 12
tmap_mode("view")
tm_shape(nga_wp)+
  tm_fill(c("total wpt","wpt functional","wpt non-functional","wpt unknown"), 
          style = "quantile", 
          palette = "Blues") +
  tm_layout(main.title = "Distribution of Water Points",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.25, 
            legend.width = 0.35,
            legend.position = c("right", "bottom"),
            frame = TRUE) +
  tm_borders(alpha = 0.5)
```
