---
title: "Birds of a feather flock together: Do functionally similar water points positively co-occur in Nigeria?"
editor: visual
execute: 
  warning: false
  message: false
---

# Overview

In this exercise, we will analyse whether the functional and non functional water points positively co-occur in Nigeria?

# Getting Started

In the code chunk below, `p_load()` of *pacman* package is used to install and load the following R packages into R environment:

-   sf,

-   tidyverse,

-   tmap,

-   spdep, and

-   funModeling will be used for rapid Exploratory Data Analysis

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse,funModeling)
```

# Importing Geospatial Data

## Importing water point geospatial

```{r}
#| eval: false

wp <- st_read(dsn = "Geodata",
              layer = "geo_export",
              crs = 4326) %>%
  filter(clean_coun == "Nigeria")
```

Things to learn from the code chunk above:

**st_read()** of sf package is used to import geo_export shapefile into R environment and save the imported geospatial data into simple feature data table.

**filter()** of dplyr package is used to extract water point records of Nigeria. In the code chunk below, "write_rds()" of reader package is used to save the extracted sf table .....

Next, `write_rds()` of readr package is used to save the extracted sf data table (i.e. wp) into an output file in rds data format. The output file is called *wp_nga.rds* and it is saved in *geodata* sub-folder.

```{r}
#| eval: false
write_rds(wp, "Geodata/wp_nga.rds")
```

# Importing Nigeria LGA boundary data

Now, we are going to import the LGA boundary data into R environment by using the code chunk below.

```{r}
#| eval: false
nga <- st_read(dsn = "Geodata",
               layer= "geoBoundaries-NGA-ADM2",
               crs = 4326)
```

Thing to learn from the code chunk above.

-   `st_read()` of **sf** package is used to import *geoBoundaries-NGA-ADM2* shapefile into R environment and save the imported geospatial data into simple feature data table.

# Data Wrangling

## Recoding the NA values into string

In the code chunk below, `replace_na()` is used to recode all the *NA* values in *status_cle* field into *Unknown*.

```{r}
#| eval: false
wp_nga <- read_rds("geodata/wp_nga.rds") %>%
  mutate(status_cle = replace_na(status_cle, "Unknown"))
```

## Exploratory Data Analysis

In the code chunk below, `freq()` of **funModeling** package is used to display the distribution of *status_cle* field in *wp_nga*.

```{r}
#| eval: false
freq(data=wp_nga, 
     input = 'status_cle')
```

# Extracting Water Points

By using classes in *status_cle* field, extracting water points records.

## Extracting functional water point

Using, `filter()` of dplyr is used to select functional water points.

```{r}
#| eval: false
wpt_functional <- wp_nga %>%
  filter(status_cle %in%
           c("Functional", 
             "Functional but not in use",
             "Functional but needs repair"))
```

```{r}
#| eval: false
freq(data=wpt_functional, 
     input = 'status_cle')
```

## Extracting non-functional water points

Using, `filter()` of dplyr is used to select non-functional water points.

```{r}
#| eval: false
wpt_nonfunctional <- wp_nga %>%
  filter(status_cle %in%
           c("Abandoned/Decommissioned", 
             "Abandoned",
             "Non-Functional",
             "Non functional due to dry season",
             "Non-Functional due to dry season"))
```

```{r}
#| eval: false
freq(data=wpt_nonfunctional, 
     input = 'status_cle')
```

## Extracting water point with unknown class

Using, `filter()` of dplyr is used to select water points with unknown status.

```{r}
#| eval: false
wpt_unknown <- wp_nga %>%
  filter(status_cle == "Unknown")
```

# Performing Point-in Polygon Count

```{r}
#| eval: false
nga_wp <- nga %>% 
  mutate(`total wpt` = lengths(
    st_intersects(nga, wp_nga))) %>%
  mutate(`wpt functional` = lengths(
    st_intersects(nga, wpt_functional))) %>%
  mutate(`wpt non-functional` = lengths(
    st_intersects(nga, wpt_nonfunctional))) %>%
  mutate(`wpt unknown` = lengths(
    st_intersects(nga, wpt_unknown)))
```

# Saving the Analytical Data Table

Learnings from code chunks:

-   `mutate()` of **dplyr** package is used to derive two fields namely *pct_functional* and *pct_non-functional*.

-   to keep the file size small, `select()` of **dplyr** is used to retain only field 3,4,9,10, 18,19,20,21,22,and 23.

```{r}
#| eval: false
nga_wp <- nga_wp %>%
  mutate(pct_functional = `wpt functional`/`total wpt`) %>%
  mutate(`pct_non-functional` = `wpt non-functional`/`total wpt`) 
```

```{r}
#| eval: false
write_rds(nga_wp, "Geodata/nga_wp.rds")
```

# Visualising the spatial distribution of water points

```{r}
#| fig-width: 20
#| fig:height: 14
nga_wp <- read_rds("Geodata/nga_wp.rds")
total <- qtm(nga_wp, "total wpt")
wp_functional <- qtm(nga_wp, "wpt functional")
wp_nonfunctional <- qtm(nga_wp, "wpt non-functional")
unknown <- qtm(nga_wp, "wpt unknown")

tmap_arrange(total, wp_functional, wp_nonfunctional, unknown, asp=1, ncol=2)
```

```{r}
nga_wp <- read_rds("Geodata/nga_wp.rds")
#| fig-width: 14
#| fig:height: 12
tm_shape(nga_wp)+
  tm_fill(c("total wpt","wpt functional","wpt non-functional","wpt unknown"), 
          style = "quantile") +
  tm_layout(main.title = "Distribution of Water Points",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.25, 
            legend.width = 0.35,
            legend.position = c("right", "bottom"),
            frame = TRUE) +
  tm_borders(alpha = 0.5)
```
