---
title: "In-Class Ex1"
editor: visual
---

# Overview

In this In-Class Ex1 , we will learn how to compute spatial weights using R. Following will be the learning at the end of this module:

-   Importing spatial and csv file data: **sf** and **readr** package

-   Relational join : **dplyr** package

-   Spatial weight computation: **spdep** package

-   Calculation of spatially lagged variables: **spdep** package

# Getting Started

The code chunk below will install and load tidyverse and sf packages.

```{r}
pacman::p_load(sf,tidyverse,spdep,tmap)
```

# Getting the data into R environment

Importing polygon features

This code chunk will import ESRI shapefile into R.

## Importing Hunan Data

```{r}
hunan = st_read(dsn="data/geospatial/", 
                layer="Hunan")
```

## Importing data attribute in csv

```{r}
hunan2012= read_csv("data/aspatial/Hunan_2012.csv")
```

## Perform relational join

Update the attribute table of *hunan*'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe

```{r}
hunan <- left_join(hunan,hunan2012)
```
